### 2017-09-03 ContinuousGeneralisation_Paper/ContinuousGeneralisation_Paper_Sigspatial/ContinuousGeneralisation_Paper_Sigspatial.tex
move to the format of ACM Sigspatial

\documentclass[sigconf, authordraft]{acmart}

\graphicspath{{figures/}}

\usepackage{booktabs} % For formal tables
\usepackage{xspace}
%\usepackage[disable]{todonotes}
\usepackage[colorinlistoftodos,prependcaption]{todonotes}
\usepackage{gensymb}  % for degree sign, e.g., temperature

%Sometimes, the links are created by the bibliography, from items with URL 
%fields. A solution that helps in this case and preserves clickable links is to 
%make sure that each \bibitem is typeset without column or page breaks:
\let\oldbibitem\bibitem
\def\bibitem{\vfill\oldbibitem}

\newcommand{\e}[1]{\times 10^{#1}}
\newcommand{\fig}{Fig.~}
\newcommand{\eq}{Eq.~}
\newcommand{\sect}{Sec.~}
\newcommand{\tab}{Tab.~}
\newcommand{\figs}{Figs.~}
\newcommand{\eqs}{Eqs.~}
\newcommand{\sects}{Secs.~}
\newcommand{\tabs}{Tabs.~}
\newcommand{\dtrm}[2][]{d_{\mathrm{#2}, t\ifthenelse{\equal{#1}{}}{}{_#1}}}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El
  Paso, Texas USA} 
\acmYear{1997}
\copyrightyear{2016}

\acmPrice{15.00}

\acmSubmissionID{123-A12-B3}

\begin{document}
\title{Continuous Generalization of Buildings by Growing and 
	Merging}
%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}

\author{Dongliang Peng}
%\authornote{Dr.~Trovato insisted his name be first.}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{Chair of Computer Science I, University of W\"urzburg}
  \streetaddress{Am Hubland}
  \city{W\"urzburg} 
  \state{Germany} 
  \postcode{97074}
}
\email{dongliang.peng@uni-wuerzburg.de}

\author{Guillaume Touya}
%\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
  \institution{COGIT, IGN}
  \streetaddress{73 avenue de Paris}
  \city{Saint-Mand\'e Cedex} 
  \state{France} 
  \postcode{94165}
}
\email{guillaume.touya@ign.fr}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{D. Peng and G. Touya}

\begin{abstract}
We propose a method to continuously 
generalize buildings from a given start scale to a smaller goal scale. 
%
If several buildings are close to each other at the start 
scale, we bridge them by line segments and merge the several buildings into 
one 
building.
For each building, either an original one or a merged one, at the start 
scale, 
we generate its target shape at 
the goal scale by enlarging and simplifying. 
Then we grow the building to the target shape.
%
After finishing the growing, we can take the goal scale as the 
start scale and repeat the process.
\newline\indent
The advantages of our method are as follows. First, the buildings 
are enlarged continuously. At the same time, they are simplified. 
Second, we ensure that the distance between any pair of two buildings is 
always larger than a specified threshold.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

\keywords{ACM proceedings, \LaTeX, text tagging}

\maketitle

\input{PaperBody}

\bibliographystyle{ACM-Reference-Format}
\bibliography{References}

\end{document}