### 2017-09-07 ContinuousGeneralisation_Paper/ContinuousGeneralisation_Paper_Sigspatial/ContinuousGeneralisation_Paper_Sigspatial.tex
working on the introduction

\documentclass[sigconf, authordraft]{acmart}

\graphicspath{{figures/}}

\usepackage{booktabs} % For formal tables
\usepackage{xspace}
%\usepackage[disable]{todonotes}
\usepackage[colorinlistoftodos,prependcaption]{todonotes}
\usepackage{gensymb}  % for degree sign, e.g., temperature

%Sometimes, the links are created by the bibliography, from items with URL 
%fields. A solution that helps in this case and preserves clickable links is to 
%make sure that each \bibitem is typeset without column or page breaks:
\let\oldbibitem\bibitem
\def\bibitem{\vfill\oldbibitem}

\newcommand{\e}[1]{\times 10^{#1}}
\newcommand{\fig}{Figure~}
\newcommand{\eq}{Equation~}
\newcommand{\sect}{Section~}
\newcommand{\tab}{Table~}
\newcommand{\figs}{Figures~}
\newcommand{\eqs}{Equations~}
\newcommand{\sects}{Sections~}
\newcommand{\tabs}{Tables~}
\newcommand{\dtrm}[2][]{d_{\mathrm{#2}, t\ifthenelse{\equal{#1}{}}{}{_#1}}}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[UrbanGIS'17]{International Workshop on Smart Cities and Urban 
Analytics}{November 2017}{Redondo Beach, California USA} 
\acmYear{2017}
\copyrightyear{2017}

\acmPrice{15.00}

\acmSubmissionID{123-A12-B3}

\begin{document}
\title{Continuous Generalization of Buildings \\ by Growing and Aggregating}
%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}

\author{Dongliang Peng}
%\authornote{Dr.~Trovato insisted his name be first.}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{Chair of Computer Science I, University of W\"urzburg}
  \streetaddress{Am Hubland}
  \city{W\"urzburg} 
  \state{Germany} 
  \postcode{97074}
}
\email{dongliang.peng@uni-wuerzburg.de}

\author{Guillaume Touya}
%\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
  \institution{COGIT, IGN}
  \streetaddress{73 avenue de Paris}
  \city{Saint-Mand\'e Cedex} 
  \state{France} 
  \postcode{94165}
}
\email{guillaume.touya@ign.fr}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{D. Peng and G. Touya}

\begin{abstract}
We propose a method to continuously 
generalize buildings from a given start scale to a smaller goal scale. 
%
If several buildings are close to each other at the start 
scale, we bridge them by line segments and merge them into 
one 
building.
For each building, either an original one or a merged one, at the start 
scale, 
we generate its target shape at 
the goal scale by enlarging and simplifying. 
Then we grow all buildings to their target shapes.
%
After finishing the growing, we can take the goal scale as the 
start scale and repeat the process.
\newline\indent
The advantages of our method are as follows. First, the buildings 
are enlarged continuously. At the same time, they are simplified. 
Second, we ensure that the distance between any pair of two buildings is 
always larger than a specified threshold.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10010405.10010476.10010479</concept_id>
	<concept_desc>Applied computing~Cartography</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Applied computing~Cartography}

\keywords{multiple representation, zooming, simplification, buffer}

\maketitle

\input{PaperBody}

\bibliographystyle{ACM-Reference-Format}
\bibliography{References}

\end{document}