### 2017-09-11 ContinuousGeneralisation_Paper/ContinuousGeneralisation_Paper_Sigspatial/ContinuousGeneralisation_Paper_Sigspatial.tex
polished abstract and introduction

\documentclass[sigconf, authordraft]{acmart}

\graphicspath{{figures/}}

\usepackage{booktabs} % For formal tables
\usepackage{xspace}
%\usepackage[disable]{todonotes}
\usepackage[colorinlistoftodos,prependcaption]{todonotes}
\usepackage{gensymb}  % for degree sign, e.g., temperature

%Sometimes, the links are created by the bibliography, from items with URL 
%fields. A solution that helps in this case and preserves clickable links is to 
%make sure that each \bibitem is typeset without column or page breaks:
\let\oldbibitem\bibitem
\def\bibitem{\vfill\oldbibitem}

\newcommand{\e}[1]{\times 10^{#1}}
\newcommand{\fig}{Figure~}
\newcommand{\eq}{Equation~}
\newcommand{\sect}{Section~}
\newcommand{\tab}{Table~}
\newcommand{\figs}{Figures~}
\newcommand{\eqs}{Equations~}
\newcommand{\sects}{Sections~}
\newcommand{\tabs}{Tables~}
\newcommand{\dtrm}[2][]{d_{\mathrm{#2}, t\ifthenelse{\equal{#1}{}}{}{_#1}}}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[UrbanGIS'17]{International Workshop on Smart Cities and Urban 
Analytics}{November 2017}{Redondo Beach, California USA} 
\acmYear{2017}
\copyrightyear{2017}

\acmPrice{15.00}

\acmSubmissionID{123-A12-B3}

\begin{document}
\title{Continuous Generalization of Buildings \\ by Growing and Aggregating}
%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}

\author{Dongliang Peng}
%\authornote{Dr.~Trovato insisted his name be first.}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{Chair of Computer Science I, University of W\"urzburg}
  \streetaddress{Am Hubland}
  \city{W\"urzburg} 
  \state{Germany} 
  \postcode{97074}
}
\email{dongliang.peng@uni-wuerzburg.de}

\author{Guillaume Touya}
%\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
  \institution{COGIT, IGN}
  \streetaddress{73 avenue de Paris}
  \city{Saint-Mand\'e Cedex} 
  \state{France} 
  \postcode{94165}
}
\email{guillaume.touya@ign.fr}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{D. Peng and G. Touya}

\begin{abstract}
We propose a method to continuously 
generalize buildings from a given start map to a smaller-scale goal map.
For the sake of convenience, 
we name the buildings on the start map \emph{original buildings}.
%
For an intermediate scale, 
we aggregate the original buildings that will be too close 
by introducing bridges.
We grow bridged original buildings based on buffering 
and simplify the grown buildings.
We take into account the shapes of the buildings 
both at the previous map and goal map to make sure that 
the buildings are always growing.
%
\newline\indent
The advantages of our method are as follows. 
First, the buildings grow continuously 
and, at the same time, are simplified.
Second, right angles of buildings are preserved during growing. 
Third, the distances between buildings are 
always larger than a specified threshold.
We do a case study to show the performances of our method.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10010405.10010476.10010479</concept_id>
	<concept_desc>Applied computing~Cartography</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Applied computing~Cartography}

\keywords{multiple representation, zooming, simplification, buffer}

\maketitle

\input{PaperBody}

\bibliographystyle{ACM-Reference-Format}
\bibliography{References}

\end{document}