### 2017-10-17 ContinuousGeneralisation_Paper/ContinuousGeneralisation_Paper_Sigspatial/ContinuousGeneralisation_Paper_Sigspatial.tex
added file about copyright

\documentclass[
sigconf,
screen, %hyperlinks are colored
%authorversion,
%authordraft %Whether author’s draft mode is enabled
]{acmart}

\graphicspath{{figures/}}

\usepackage{booktabs} % For formal tables
\usepackage{xspace}
%\usepackage[disable]{todonotes}
\usepackage[colorinlistoftodos,prependcaption,disable]{todonotes}
\usepackage{gensymb}  % for degree sign, e.g., temperature
\usepackage{balance}
%\usepackage{tikz,pgfplots}
%\pgfplotsset{compat=1.3} %make y lable compact in a tikz 
%%picture with axes

%Sometimes, the links are created by the bibliography, from items with URL 
%fields. A solution that helps in this case and preserves clickable links is to 
%make sure that each \bibitem is typeset without column or page breaks:
\let\oldbibitem\bibitem
\def\bibitem{\vfill\oldbibitem}

\newcommand{\fig}{Figure~}
\newcommand{\eq}{Equation~}
\newcommand{\sect}{Section~}
\newcommand{\tab}{Table~}
\newcommand{\figs}{Figures~}
\newcommand{\eqs}{Equations~}
\newcommand{\sects}{Sections~}
\newcommand{\tabs}{Tables~}
\newcommand{\dtrm}[2][]{d_{\mathrm{#2}, 
t\ifthenelse{\equal{#1}{}}{}{_#1}}}

\begin{document}
\title{Continuously Generalizing Buildings to Built-up Areas\\ 
by Aggregating and Growing}
%\titlenote{Produces the permission block, and
%  copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}

\author{Dongliang Peng}
%\authornote{Dr.~Trovato insisted his name be first.}
\orcid{0000-0001-6848-3545}
\affiliation{%
  \institution{Chair of Computer Science I, University of W\"urzburg}
  \streetaddress{Am Hubland}
  \city{W\"urzburg} 
  \state{Germany} 
  \postcode{97074}
}
\email{dongliang.peng@uni-wuerzburg.de}

\author{Guillaume Touya}
%\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
  \institution{COGIT, IGN}
  \streetaddress{73 avenue de Paris}
  \city{Saint-Mand\'e Cedex} 
  \state{France} 
  \postcode{94165}
}
\email{guillaume.touya@ign.fr}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{D. Peng and G. Touya}

\begin{abstract}
To enable smooth zooming, we propose a method to continuously 
generalize buildings from a given start map to a smaller-scale goal map, where 
there are only built-up area polygons instead of individual building polygons.
We name the buildings on the start map \emph{original buildings}.
%
For an intermediate scale, 
we aggregate the original buildings that will become too close 
by adding bridges.
We grow (bridge) original buildings based on buffering, 
and we also simplify the grown buildings.
We take into account the shapes of the buildings 
both at the previous map and goal map to make sure that 
the buildings are always growing.
%
\newline\indent
The advantages of our method are as follows. 
First, the buildings grow continuously 
and, at the same time, are simplified.
Second, right angles of buildings are preserved during growing: the merged 
buildings still look like buildings. 
Third, the distances between buildings are 
always larger than a specified threshold.
We do a case study to show the performances of our method.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10010405.10010476.10010479</concept_id>
	<concept_desc>Applied computing~Cartography</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Applied computing~Cartography}

\copyrightyear{2017}
\acmYear{2017}
\setcopyright{othergov}
\acmConference[UrbanGIS'17]{UrbanGIS'17:3rd ACM SIGSPATIAL 
	Workshop on Smart Cities and Urban Analytics }{November 
	7--10, 
	2017}{Redondo Beach, CA, USA}
\acmBooktitle{UrbanGIS'17: UrbanGIS'17:3rd ACM SIGSPATIAL 
	Workshop on Smart Cities and Urban Analytics , November 
	7--10, 
	2017, Redondo Beach, CA, USA}
\acmPrice{15.00}
\acmDOI{10.1145/3152178.3152188}
\acmISBN{978-1-4503-5495-0/17/11}

\keywords{multiple representation, zooming, simplification, buffer}

\maketitle

\input{PaperBody}

\bibliographystyle{ACM-Reference-Format}
\balance %balance the two columns of references at the last page
\bibliography{References}

\end{document}